Dominance Regions for Affine Cluster Algebras

### 2022-06-19 affine_dominance.tex
Mutation matrices for g-vectors along the preprojective seeds

  Let $B=(b_{ij})$ denote an acyclic $n\times n$ exchange matrix.
  Write $Q_B$ for the corresponding quiver, here $b_{ij}>0$ means there is an arrow from $j$ to $i$ in $Q$.
  Without loss of generality we assume $1,\ldots,n$ is a sink adapted sequence for $Q_B$, i.e. vertex $1$ is a sink and vertex $n$ is a source in $Q$.
  This implies $B$ has positive entries above the diagonal.
  \sayDR{This seemed to give the cleanest calculation.}
  Define vectors $\bfc^{(k)}_r\in\ZZ^n$ for $1\le k\le n$ and $r\in\ZZ_{\ge0}$ recursively as follows:
  \sayDR{A slight modification is needed to handle the opposite mutation sequence.}
  \begin{itemize}
    \item $\bfc^{(k)}_0=\bfe_k$ is the standard basis vector 
    \item for $1\le \ell \le n$, we have $\bfc^{(k)}_{ns+\ell}=\begin{cases} -\bfc^{(k)}_{ns+\ell-1} & \text{if $\ell=k$;}\\ \bfc^{(k)}_{ns+\ell-1}+b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1} & \text{if $\ell<k$;}\\ \bfc^{(k)}_{ns+\ell-1}-b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1} & \text{if $\ell>k$.} \end{cases}$
  \end{itemize}
  For $r\in\ZZ$ and $1\le k\le n$, write $r\equiv_n k$ if $n|(r-k)$ and $r\not\equiv_n k$ otherwise.
  \begin{lemma}
    \label{lem:tropical signs}
    Assume $B$ is not finite-type.
    For $r\in\ZZ_{\ge0}$, the vector $\bfc^{(k)}_r$ is non-negative if $r\not\equiv_n k$ and non-positive when $r\equiv_n k$.
  \end{lemma}
  \begin{proof}
    Induction?
  \end{proof}

  \begin{lemma}
    For any $s\in\ZZ_{\ge0}$ and $0\le \ell \le n-1$, the matrix $\mu_\ell\cdots\mu_1(\mu_n\cdots\mu_1)^s\tilde B$ has principal part given by $(b'_{ij})$ where $b'_{ij}=(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}$ for $\epsilon(i,j)=\begin{cases} 1 & \text{if $i\le j$;}\\ 0 & \text{otherwise;}\end{cases}$ and the $\bfc$-vector part of the $k$-th column is given by $\bfc^{(k)}_{ns+\ell}$ for $1\le k\le n$.
  \end{lemma}
  \begin{proof}
    We work by induction on $ns+\ell$.
    For $s=\ell=0$, we just get $\tilde B$ and the claim clearly holds.
    Assume the claim holds for a chosen $s$ and $\ell-1$, we show the claim holds for $s$ and $\ell$ (the case $\ell=n$ gives rise to the claim for $s+1$ and $\ell=0$).
    Write $\tilde B'=\mu_{\ell-1}\cdots\mu_1(\mu_n\cdots\mu_1)^s\tilde B$ and let $Q'_B$ denote the corresponding acyclic quiver.

    By \eqref{eq:matrix mutation}, the mutation $\mu_\ell$ reverses the sign of the $\ell$-th row and column of $\tilde B'$.
    It follows that the principal part of $\mu_\ell\tilde B'$ is given as follows:
    \begin{itemize}
      \item if $i,j<\ell$, or $i<\ell$ and $j>\ell$, or $i<\ell$ and $j>\ell$, then $(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}=(-1)^{\epsilon(i,\ell-1)+\epsilon(j,\ell-1)}b_{ij}$;
      \item if $i=\ell$ or $j=\ell$, then $(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}=-(-1)^{\epsilon(i,\ell-1)+\epsilon(j,\ell-1)}b_{ij}$;
    \end{itemize}
    in each case we obtain the desired expression.
    In addition, we see that the $\ell$-th $\bfc$-vector is given by $-\bfc^{(\ell)}_{ns+\ell-1}=\bfc^{(\ell)}_{ns+\ell}$ as desired.
    By Lemma~\ref{lem:tropical signs}, the $\ell$-th $\bfc$-vector of $\tilde B'$ is non-negative.  
    Observe further that vertex $\ell$ is a sink in $Q'_B$ so that the $\ell$-th row of $\tilde B'$ has non-negative entries, in particular we see that $b'_{ij}=-b_{ij}$ for $j<\ell$ and $b'_{ij}=b_{ij}$ for $j>\ell$.
    It then follows from \eqref{eq:matrix mutation} that the $k$-th $\bfc$-vector of $\mu_\ell\tilde B'$ is given by $\bfc^{(k)}_{ns+\ell-1}+b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1}$ if $\ell<k$ and by $\bfc^{(k)}_{ns+\ell-1}-b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1}$ if $\ell>k$.
    Each of these is equal to $\bfc^{(k)}_{ns+\ell}$ as desired.
  \end{proof}

>>>>

  Let $B=(b_{ij})$ denote an acyclic $n\times n$ exchange matrix.
  Write $Q_B$ for the corresponding quiver, here $b_{ij}>0$ means there is an arrow from $j$ to $i$ in $Q$.
  \sayDR{Note that $Q_B$ is only recording the orientations of edges.}
  Without loss of generality we assume $1,\ldots,n$ is a sink adapted sequence for $Q_B$, i.e. vertex $1$ is a sink and vertex $n$ is a source in $Q$.
  This implies $B$ has positive entries above the diagonal.
  \sayDR{This seemed to give the cleanest calculation.}
  Define vectors $\bfc^{(k)}_r\in\ZZ^n$ for $1\le k\le n$ and $r\in\ZZ_{\ge0}$ recursively as follows:
  \sayDR{A slight modification is needed to handle the opposite mutation sequence.}
  \begin{itemize}
    \item $\bfc^{(k)}_0=\bfe_k$ is the standard basis vector 
    \item for $1\le \ell \le n$, we have $\bfc^{(k)}_{ns+\ell}=\begin{cases} -\bfc^{(k)}_{ns+\ell-1} & \text{if $\ell=k$;}\\ \bfc^{(k)}_{ns+\ell-1}+b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1} & \text{if $\ell<k$;}\\ \bfc^{(k)}_{ns+\ell-1}-b_{\ell k}\bfc^{(\ell)}_{ns+\ell-1} & \text{if $\ell>k$.} \end{cases}$
  \end{itemize}
  For $r\in\ZZ$ and $1\le k\le n$, write $r\equiv_n k$ if $n|(r-k)$ and $r\not\equiv_n k$ otherwise.
  \begin{lemma}
    \label{lem:tropical signs}
    Assume $B$ is not finite-type.
    For $r\in\ZZ_{\ge0}$, the vector $\bfc^{(k)}_r$ is non-negative if $r\not\equiv_n k$ and non-positive when $r\equiv_n k$.
  \end{lemma}
  \begin{proof}
    Induction?
  \end{proof}

  Let $\TT_n$ denote the $n$-regular labeled tree with root vertex $t_0$.
  Define vertices $t_r\in\TT_n$ for $r\in\ZZ$ by joining $t_{r-1}$ to $t_r$ by an edge labeled $\ell$ if $r\equiv_n\ell$.
  Since $1,\ldots,n$ is sink adapted, this labels preprojective seeds by positive integers and postinjective seeds by negative integers.
  Write $\tilde B^{t_r}$ for the extended exchange matrix associated to the seed at $t_r$.
  \begin{lemma}
    For $r\in\ZZ_{\ge0}$ with $r=ns+\ell$ for $1\le \ell \le n$, the matrix $\tilde B^{t_r}$ has principal part given by $b^{t_r}_{ij}=(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}$ for $\epsilon(i,j)=\begin{cases} 1 & \text{if $i\le j$;}\\ 0 & \text{otherwise;}\end{cases}$ and the $\bfc$-vector part of the $k$-th column is given by $\bfc^{(k)}_r$ for $1\le k\le n$.
  \end{lemma}
  \begin{proof}
    We work by induction on $ns+\ell$.
    For $s=\ell=0$, we just get $\tilde B$ and the claim clearly holds.
    Assume the claim holds for $r-1$, we show the claim holds for $r$.

    By \eqref{eq:matrix mutation}, the mutation $\mu_\ell$ reverses the sign of the $\ell$-th row and column of $\tilde B^{t_{r-1}}$.
    It follows that the principal part of $\tilde B^{t_r}$ is given as follows:
    \begin{itemize}
      \item if $i,j<\ell$, or $i<\ell$ and $j>\ell$, or $i<\ell$ and $j>\ell$, then 
        \[b^{t_r}_{ij}=b^{t_{r-1}}_{ij}=(-1)^{\epsilon(i,\ell-1)+\epsilon(j,\ell-1)}b_{ij}=(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij};\]
      \item if $i=\ell$ or $j=\ell$, then $(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij}=-(-1)^{\epsilon(i,\ell-1)+\epsilon(j,\ell-1)}b_{ij}$;
        \[b^{t_r}_{ij}=-b^{t_{r-1}}_{ij}=-(-1)^{\epsilon(i,\ell-1)+\epsilon(j,\ell-1)}b_{ij}=(-1)^{\epsilon(i,\ell)+\epsilon(j,\ell)}b_{ij};\]
    \end{itemize}
    in each case we obtain the desired expression.
    In addition, we see that the $\ell$-th $\bfc$-vector is given by $-\bfc^{(\ell)}_{r-1}=\bfc^{(\ell)}_r$ as desired.

    By Lemma~\ref{lem:tropical signs}, the $\ell$-th $\bfc$-vector of $\tilde B^{t_{r-1}}$ is non-negative.  
    Observe further that vertex $\ell$ being a sink implies that the $\ell$-th row of $\tilde B^{t_{r-1}}$ has non-negative entries, in particular we see that $b^{t_{r-1}}_{\ell k}=b_{\ell k}$ for $\ell<k$ and $b^{t_{r-1}}_{\ell k}=-b_{\ell k}$ for $\ell>k$.
    It then follows from \eqref{eq:matrix mutation} that the $k$-th $\bfc$-vector of $\tilde B^{t_r}$ is given by $\bfc^{(k)}_{r-1}+b_{\ell k}\bfc^{(\ell)}_{r-1}$ if $\ell<k$ and by $\bfc^{(k)}_{r-1}-b_{\ell k}\bfc^{(\ell)}_{r-1}$ if $\ell>k$.
    Each of these is equal to $\bfc^{(k)}_r$ as desired.
  \end{proof}

  It follows that we may write explicit expressions for the $2n\times 2n$ mutation matrices $E^{t_r}_{k,\varepsilon}$ for $r\ge0$.
  Indeed, taking $r=ns+\ell$ with $1 \le \ell \le n$ we get $E^{t_r}_{k,\varepsilon}=(e^{t_r}_{ij})$ with
  \begin{equation}
    \label{eq:preprojective mutation matrices}
    e^{t_r}_{ij}=\begin{cases} 1 & \text{if $i=j\ne k$;}\\ -1 & \text{if $i=j=k$;}\\ b_{ik} & \text{if $\varepsilon=+$ and $i<j=k\le\ell$ or $j=k\le\ell<i$ or $\ell<i<j=k$;}\\ -b_{ik} & \text{if $\varepsilon=-$ and $j=k<i\le\ell$ or $i\le\ell<j=k$ or $\ell<j=k<i$;} \\ 0 & \text{otherwise;} \end{cases}
  \end{equation}

  \[
    \left[\begin{array}{cc|c|cccc}
      0 & + & + & - & - & \cdots & -\\
      - & 0 & + & - & - & \cdots & -\\
      \hline
      - & - & 0 & - & - & \cdots & -\\
      \hline
      + & + & + & 0 & + & \cdots & +\\
      + & + & + & - & 0 & \cdots & +\\
      \vdots & \vdots & \vdots & \vdots & \vdots & \ddots & \vdots\\
      + & + & + & - & - & \cdots & 0
    \end{array}\right]
  \]
  where the bottom $n$ entries in column $k$ are given as follows:
  \begin{itemize}
    \item if $\varepsilon=+$, we have $\bfc^{(k)}_r$ if $r\not\equiv_n k$ and zero otherwise;
    \item if $\varepsilon=-$, we have $-\bfc^{(k)}_r$ if $r\equiv_n k$ and zero otherwise.
  \end{itemize}
